---
title: "Analysis"
author: "Ben"
date: "7/11/2018"
output: html_document
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
```

# Simulating UID negative correlation with UID cost in speaker/listener populations

### Summary

Simply interacting with other agents gives rise to the effect in a population of speakers -- alternative take on fixed point analysis....


```{r read-data}
fp_rsa_100 <- "../outputs/agent-100-full.csv"           # not permuted conversation partners
fp_no_rsa_100 <- "../outputs/agent-100-no-rsa-full.csv" # not permuted conversation partners
fp_rsa_10 <- "../outputs/agent-10-full.csv"
fp_no_rsa_10 <- "../outputs/agent-10-no-rsa-full.csv"
fp_rsa_6 <- "../outputs/agent-6-full.csv"
fp_no_rsa_6 <- "../outputs/agent-6-no-rsa-full.csv"
fp_rsa_4 <- "../outputs/agent-4-full.csv"
fp_no_rsa_4 <- "../outputs/agent-4-no-rsa-full.csv"
fp_rsa_2 <- "../outputs/agent-2-full.csv"
fp_no_rsa_2 <- "../outputs/agent-2-no-rsa-full.csv"
fp_rsa_1 <- "../outputs/agent-1-full.csv"
fp_no_rsa_1 <- "../outputs/agent-1-no-rsa-full.csv"

df <- rbind(
  # read.csv(fp_rsa_100) %>% mutate(use_rsa=TRUE),
  # read.csv(fp_no_rsa_100) %>% mutate(use_rsa=FALSE),
  read.csv(fp_rsa_10) %>% mutate(use_rsa=TRUE),
  read.csv(fp_no_rsa_10) %>% mutate(use_rsa=FALSE),
  read.csv(fp_rsa_6) %>% mutate(use_rsa=TRUE),
  read.csv(fp_no_rsa_6) %>% mutate(use_rsa=FALSE),
  read.csv(fp_rsa_4) %>% mutate(use_rsa=TRUE),
  read.csv(fp_no_rsa_4) %>% mutate(use_rsa=FALSE),
  read.csv(fp_rsa_2) %>% mutate(use_rsa=TRUE),
  read.csv(fp_no_rsa_2) %>% mutate(use_rsa=FALSE),
  read.csv(fp_rsa_1) %>% mutate(use_rsa=TRUE),
  read.csv(fp_no_rsa_1) %>% mutate(use_rsa=FALSE)
)
```

```{r preprocess-1}
df_pop <- df %>%
  filter(id=='pop')
df_no_pop <- df %>%
  filter(id!='pop')
```

Preprocessing
```{r preprocess-get-max-round}
df_stable <- df_pop[!is.na(df_pop$r), ]

# Get max rounds by id, sim_id
max_rounds <- df_stable %>%
  group_by(use_rsa, n_agents, id, sim_id) %>%
  summarise(max_round=max(round))
get_max_round <- function(use_rsa_, n_agents_, id_, sim_id_) {
  subset(max_rounds, (use_rsa==use_rsa_ & n_agents==n_agents_ & id==id_ & sim_id==sim_id_))$max_round
}
# Attempt with mapply -- note this is slow running with we don't restrict to id==pop (for larger speaker populations).
max_rounds_list <- mapply(get_max_round, df_stable$use_rsa, df_stable$n_agents, df_stable$id, df_stable$sim_id)
# Attempt with purrr:pmap
# max_rounds_list <- df_stable %>%
#   select(use_rsa, n_agents, id, sim_id) %>%
#   purrr::pmap(., function(use_rsa, n_agents, id, sim_id) {get_max_round(use_rsa, n_agents, id, sim_id)}) %>%
#   unlist()
assertthat::are_equal(length(max_rounds_list), nrow(df_stable))

# Assign max round
df_stable$max_round <- max_rounds_list
df_stable <- df_stable %>%
  mutate(stable=max_round==n_rounds-1,
         is_final_round=max_round==round)
```

### Samples

Note the facet numbers are not meaningful... should show this another way.
```{r sample-that-rate-trajectories}
sims <- sample(unique(df_stable$sim_id), min(length(unique(df_stable$sim_id)), 16))

df_stable %>%
  filter(sim_id %in% sims, id=='pop') %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=round, y=that_rate, lty=use_rsa, col=as.factor(n_agents))) +
    geom_line(alpha=0.9) +
    ylim(0, 1) +
    ggtitle("Sample that-rate trajectories") +
    facet_wrap(~sim_id) +
    theme_classic()
```

```{r sample-r-trajectories}
df_stable %>%
  filter(sim_id %in% sims, id=='pop') %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=round, y=r, lty=use_rsa, col=as.factor(n_agents))) +
    geom_line(alpha=0.9) +
    ylim(-1, 1) +
    ggtitle("Sample r trajectories") +
    facet_wrap(~sim_id) +
    theme_classic() 
```

### `r` Trajectories

```{r}
df_stable %>%
  filter(id=='pop') %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=round, y=r, lty=use_rsa, col=as.factor(sim_id))) +
    geom_line(alpha=0.6) +
    ylim(-1, 1) +
    theme_classic() +
    theme(legend.position='none') +
    facet_wrap(n_agents~use_rsa, nrow=5)
```

### `that-rate` Trajectories
```{r all-population-that-rate-trajectories}
df_stable %>%
  filter(id=='pop') %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=round, y=that_rate, col=as.factor(sim_id))) +
    geom_line(alpha=0.6) +
    ylim(0, 1) +
    theme_classic() +
    theme(legend.position='none') +
    facet_wrap(n_agents~use_rsa, nrow=5)
```

### `that-rate` distr
```{r}
df_stable %>%
  filter(is_final_round, id=='pop') %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=that_rate,y=..density..)) + 
      geom_histogram(bins=32) + 
      scale_x_continuous(limits=c(-0.05,1.05)) +
      ylab("Probability density") +
      xlab(expression(paste("Marginal frequency of optional marker ", t))) +
      theme_classic() +
      facet_wrap(n_agents~use_rsa, nrow=5)
```

### `r` distr
```{r}
df_stable %>% 
  filter(is_final_round, id=='pop') %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=r, y=..density.., fill=as.factor(n_agents))) + 
    geom_vline(xintercept=0.0, lty=2, alpha=0.3) +
    geom_histogram(bins=42, alpha=0.9) + 
    scale_x_continuous(limits=c(-1.05,1.05)) +
    ylab("Probability density") +
    xlab("Pearson correlation between\nphrase onset & t probabilities") + 
    theme_classic() +
    facet_grid(n_agents~use_rsa)
```

### Significant diffs?

this is actually really interesting -- the effect is not a fun
```{r means-plot}
df_stable %>%
  filter(is_final_round) %>%
  group_by(use_rsa, n_agents) %>%
  summarise(mean_r=mean(r),
            sd_r=sd(r),
            n=n(),
            ci_hi=mean_r+qnorm(0.975)*sd_r/sqrt(n),
            ci_lo=mean_r+qnorm(0.025)*sd_r/sqrt(n)) %>%
  ggplot(aes(x=as.factor(n_agents), y=mean_r, fill=use_rsa)) +
    geom_bar(stat='identity', position='dodge') +
    geom_errorbar(aes(ymin=ci_lo, ymax=ci_hi), width=0.2, position=position_dodge(0.9)) +
    ylim(-0.3, 0.1) +
    theme_classic()
```
```{r plo1-means-2}
df_stable %>%
  filter(is_final_round) %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=as.factor(n_agents), y=r, col=use_rsa, fill=use_rsa)) +
    geom_hline(yintercept=0.0, lty=2, alpha=0.2) +
    geom_point(aes(col=use_rsa), alpha=0.4) +
    # geom_boxplot(alpha=0) +
    geom_violin(alpha=0.3) +
    
    coord_flip() +
    theme_classic() +
    facet_wrap(~use_rsa)
```


## Convergence and entropy

Average lexicon entropy decreases across RSA conversations.
```{r plot-lexicon-entropy}
df_pop %>%
  group_by(use_rsa, n_agents, round) %>%
  summarise(mean_entropy=mean(entropy),
            sd_entropy=sd(entropy),
            n=n(),
            ci_hi=mean_entropy+qnorm(0.975)/sqrt(n),
            ci_lo=mean_entropy+qnorm(0.025)/sqrt(n)) %>%
  ungroup() %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=round, y=mean_entropy, col=use_rsa)) +
    geom_line() +
    geom_errorbar(aes(ymin=ci_lo, ymax=ci_hi), alpha=0.3) +
    theme_classic() +
    facet_wrap(~n_agents)
```

```{r}
df_stable %>%
  filter(is_final_round) %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=as.factor(n_agents), y=entropy, fill=use_rsa)) +
    geom_boxplot() +
    theme_classic()
```

Decerease in lexicon entropy.
```{r}
df_stable %>%
  filter(is_final_round, id=='pop')  %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  select(use_rsa, n_agents, sim_id, entropy) %>%
  spread(use_rsa, entropy) %>%
  mutate(diff=rsa-`no-rsa`) %>%
  group_by(n_agents) %>%
  ggplot(aes(x=as.factor(n_agents), y=diff, fill=as.factor(n_agents))) +
    geom_hline(yintercept=0, lty=2, alpha=0.4, col='red')  +
    geom_violin(aes(fill=as.factor(n_agents)), alpha=0.4, trim=FALSE) +
    geom_point(alpha=0.5) +
    coord_flip() +
    theme_classic()
```


Due to lower overall entropy individual speakers lexicons converge more closely, perhaps as a function of populations size.
```{r plot-avg-speaker-kl}
df_no_pop %>% 
  group_by(use_rsa, n_agents, round) %>%
  summarise(mean_KL=mean(KL),
            sd_KL=sd(KL),
            n=n(),
            ci_hi=mean_KL+qnorm(0.975)/sqrt(n),
            ci_lo=mean_KL+qnorm(0.025)/sqrt(n)) %>%
  ungroup() %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  ggplot(aes(x=round, y=mean_KL, col=as.factor(n_agents))) +
    geom_line() +
    # geom_errorbar(aes(ymin=ci_lo, ymax=ci_hi)) +
    ylim(0, 0.3) +
    theme_classic() +
    facet_wrap(~use_rsa)
```

B-prob example (one simulation). What should aggregate statistic be?

The population converges on a shared distribution over relative clauses (worth thinking as need-probabilities).
```{r}
df_no_pop %>%
  filter(sim_id %in% c(10)) %>%
  ggplot(aes(x=round, y=current_B_prob, col=as.factor(paste0(id, '-', sim_id)))) +
    geom_line() +
    theme_classic() +
    facet_grid(use_rsa~n_agents)
```

t-prob example (one simulation). What should aggregate statistic be?

However, there is *still* lots of individual-level variation in terms of optinal marking for RSA agents...
```{r}
df_no_pop %>%
  filter(sim_id %in% c(10)) %>%
  ggplot(aes(x=round, y=current_t_prob, col=as.factor(paste0(id, '-', sim_id)))) +
    geom_line() +
    theme_classic() +
    facet_grid(use_rsa~n_agents)
```

Variance within a population `optional-t` usage increases for RSA agents.

How to interpret this? As the conversations proceed the within population variance of `optional-t` usage increases for RSA speakers and declines for non-rsa speakers.
```{r}
df_no_pop %>%
  mutate(use_rsa=factor(ifelse(use_rsa, 'rsa', 'no-rsa'), levels=c('rsa', 'no-rsa'))) %>%
  group_by(use_rsa, n_agents, id, round) %>%
  summarise(mean_current_t_probs=mean(current_t_prob),
            sd_current_t_probs=sd(current_t_prob)^2) %>%
  group_by(use_rsa, n_agents, round) %>%
  summarise(mean_sd_current_t_probs=mean(sd_current_t_probs),
            sd_sd_current_t_probs=sd(sd_current_t_probs),
            n=n(),
            ci_lo=mean_sd_current_t_probs+qnorm(0.025)*sd_sd_current_t_probs/sqrt(n),
            ci_hi=mean_sd_current_t_probs+qnorm(0.975)*sd_sd_current_t_probs/sqrt(n)) %>%
  ggplot(aes(x=round, y=mean_sd_current_t_probs, col=as.factor(n_agents))) +
    geom_errorbar(aes(ymin=ci_lo, ymax=ci_hi), alpha=0.4) +
    geom_line() +
    theme_classic() +
    facet_grid(use_rsa~., scales="free")
```



```{r}
summary(lm(r~as.factor(n_agents)+use_rsa, data=df_stable))
```


### Stable optionality plot

```{r eval=FALSE}
# R. Levy's preprocessing
# dat$stable <- with(dat,thatrate > 0.001 & thatrate < 0.999)
# dat <- subset(dat, ! (k==1.0 & c==0.0))
# dat.summary <- dat %>% group_by(k,c) %>%
#   dplyr:::summarise(stable=mean(stable),r=mean(r))

# This preprocessing is only valid for stable optionality plot
df_preprocessed <- df %>%
  filter(al_round, id!='pop') %>%
  filter(k != 1.0, c != 0.0) %>%
  group_by(id, k, c) %>%
  summarise(stable=mean(is_stable), r=mean(r))
```


```{r eval=FALSE}
stables <- df %>% 
  filter(is_final_round) %>%
  mutate(stable=ifelse(that_rate > 0.1 & that_rate < 0.9, TRUE, FALSE)) %>%
  filter(stable)
stables <- stables$sim_id
# Plot all trajectories  
```

```{r plot-reproduce-stable-optionality-grid, eval=FALSE}
df_preprocessed %>%
  filter(id!='pop') %>%
  ggplot(aes(k,c)) + 
    geom_tile(aes(fill=r),colour="white") +
    labs(y=expression(paste("String length cost parameter ", c)), fill="stable\noptionality\nrate") +
    theme_classic() +
    scale_x_continuous(name=expression(paste("Nonuniformity penalization parameter ",k)),
                       breaks=seq(0,2,by=0.2))
```



